---
title: "Code Chunks"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Updating Coverage

If you want to add to the current coverage without having to restate all the other coverage elements using `set_coverage`, these chunks allow you to just add taxonomic coverage and single date temporal coverage. 

### Adding taxonomic coverage

```{r, eval = FALSE}
# add each new element as a tax object
tax1 <- new("taxonomicClassification",
            taxonRankName = new("taxonRankName", "Species"),
            taxonRankValue = new("taxonRankValue", "Calamagrostis deschampsioides"))

tax2 <- new("taxonomicClassification", 
            taxonRankName = new("taxonRankName", "Species"),
            taxonRankValue = new("taxonRankValue", "Carex aquatilis"))

# combine all tax elements into taxonomic coverage object
taxcov <- new("taxonomicCoverage",
               taxonomicClassification = c(tax1, tax2))

eml@dataset@coverage@taxonomicCoverage <- c(taxcov)

```

### Adding single date temporal coverage

```{r, eval = FALSE}
date <- new("singleDateTime",
           calendarDate = "2011-09-15")

tempcov1 <- new("temporalCoverage",
               singleDateTime = date)

eml@dataset@coverage@temporalCoverage <- c(tempcov1)
```

## Data Objects
### Adding data tables for a whole folder of files with the same attributes

```{r, eval = FALSE}
# set path to data folder
data_path <- '/home/visitor/…”' 

# list the files. recursive = TRUE will get all files even if they are in folders in folders
paths <- list.files(data_path, full.names=TRUE, recursive = TRUE) 

# generate pids. replace format_id with the one that applies
new_pids1 <- sapply(paths, function(path) {publish_object(mn, path, format_id = "text/plain")}) 

attributes1 <- read.csv("/home/me/file_attributes.csv", stringsAsFactors = FALSE, na.strings = c(''))
attributeList1 <- set_attributes(attributes1)

# assign attributes to data tables
eml <- eml_add_entities(eml, data.frame(type = "dataTable",
                                        path = paths,
                                        pid = new_pids1,
                                        format_id = "text/plain"))
eml@dataset@dataTable[[1]]@attributeList <- attributeList1
eml@dataset@dataTable[[1]]@attributeList@id <- new("xml_attribute", "shared_attributes1")

#run a loop over all the data pids to add data tables with descriptions
for (i in 2:length(eml@dataset@dataTable)) {
  eml@dataset@dataTable[[i]]@attributeList@references <- new("references", "shared_attributes1")
  eml@dataset@dataTable[[i]]@entityDescription <- new("entityDescription", .Data = "entity description")
}

```

## System Metadata
### Obsolescence chain

This chunk is to obsolete one dataset. If there are more to add to the chain, more steps can be added, carefully making sure to fill in `obsoletes` and `obsoletedBy` slots for each one. 

```{r, eval = FALSE}
# get oldest version of the file you want to be visible. Use get_all_versions and look at the first.
# urn:uuid:...

# PID for data set to be hidden: doi:10…

# adding data set to hide in the slot before the first version of the visible data set
gsmOG <- getSystemMetadata(mn, "urn:uuid:...")
gsmOG@obsoletes <- "doi:10…"
updateSystemMetadata(mn, "urn:uuid:...", gsmOG)

# adding first version to obsolescence chain after hidden version
gsmObs1 <- getSystemMetadata(mn, "doi:10…")
gsmObs1@obsoletedBy <- "urn:uuid:..."
updateSystemMetadata(mn, "doi:10…", gsmObs1)

```

### Set rights and access

This chunk sets rights and access for metadata, resource map, and all data objects in a package

```{r, eval = FALSE}
## Fix rights holder and access
PI_name <- "http://orcid.org/…."
project_pid <- "resource_map_doi:10…."
project_package <- get_package(mn, project_pid)
set_rights_and_access(mn, c(project_package$metadata, project_package$resource_map, project_package$data), PI_name, c("read", "write", "changePermission"))

```

## Miscellaneous 

### Adding sampling info to methods section

```{r, eval = FALSE}
step1 <- new('methodStep',
             description = "methods text")

stEx <- new("studyExtent", 
           description = "study extent description")

samp <- new("sampling", 
           studyExtent = stEx,
           samplingDescription = "sampling description text")

methods1 <- new("methods",
               methodStep = c(step1),
               sampling = samp)
eml@dataset@methods <- methods1
```

### Adding a pre generated identifier to the eml

When you pre generate a UUID or DOI, the change is not automatically reflected in the "Identifier" section of the eml so this makes sure that the eml lines up with the identifier being used. 

```{r, eval = FALSE}
## Generate DOI and add to EML
doiPid <- generateIdentifier(mn, "DOI")
eml@packageId <- new("xml_attribute",
                      .Data = doiPid)

```

### Dealing with netCDFs

This section is for dealing with netCDF (.nc) files. These files require data tables but since they can not be simply opened on the computer using a text editor or Excel, you can use Panoply to explore them or these R commands:


```{r}
library(arcticdatautils)
library(ncdf4)
filepath <- '/home/sharisnochs/Petterson/ICECAPS_precip_product_2010_2015.nc' 

# gets attribute info
atts <- get_ncdf4_attributes(filepath)
# preview of View(atts)
atts[1:5,]


# returns the actual values for a specified attribute
t <- nc_open(filepath)
test <- ncvar_get(t, 't2m')
# preview of View(test)
test[1:10]
```

The `formatId` in the sysmeta will most likely be `netCDF-4`.