---
title: "Frequently Used Code Chunks"
author: "Vivian Tran"
date: "3/7/2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

These are some code chunks that I frequently come back to when processing data for the Arctic Data Center. 

#Reading in raw data
##Single data file
```{r eval=FALSE}
df <- read.table("path/to/data", 
               header=T, 
               fill = T, # blank fields are added for rows that have unequal length,
               sep= ",", # put "," for .csv file, "\t" for files with values separated by tabs
               na.strings = c('','NA')) # fills blank rows with NA's 
```
Some things to note: I specify fill=T because I often get an error related to the lengths of rows/columns:

```{r echo=FALSE}
cat("Error in scan(file = file, what = what, sep = sep, quote = quote, dec = dec,  : 
  line 1 did not have 10 elements")
```

This usually happens in .txt files.

  
##Multiple data files 
I use these chunks when I want to read in several data files that share the same column names and formatting. I usually group my data files into different folders according to the type of data and/or formatting to facilitate reading in the data. This will help automate reformatting later on. 
```{r eval=FALSE}
# grab data paths from the folder that data is stored in 
# "path" specifies name of folder that data paths are stored 
# full.names = T produces full paths to fukes instead of just file name

rawPaths <- dir(path = "path/to/folder", full.names = T) 
```

Read in data using a for loop. Remember to initialize all variables that you will be using outside of the for loop. 
```{r eval=FALSE}
dataList <- vector("list", length(rawPaths)) # makes an empty list with same length as file paths vector
i=0 
for(i in 1:length(rawPaths)){
  dataList[[i]] <- read.table(rawPaths[i], 
                               na.strings = c("", "NA"),
                               header=T)  
}
```
Note: list() creates an empty list of length 0. However, vector("list", length(rawPaths)) allocates a designated number of slots within the list instead of the list being constantly updated every time the for loop interates. With a small number of iterations, the time it takes for the code to run is not noticeable. However, for a large number of iterations, not allocating space will cause the code to run very slowly. 


#Removing Extraneous Rows and Columns

##Rows

Iterate through all the rows in a data frame. 
allRows is a vector containing "TRUE" and "FALSE". Each element corresponds to a row in dataFrame. 
is.na(dataFrame[i,]) outputs "TRUE" if the row contains at least one blank cell, and "FALSE" otherwise. 
all(is.na(dataFrame[i,])) outputs "TRUE" if all cells in that row are blank, and "FALSE" otherwise. 
```{r eval=F}
i=0
allRows <- c() # initialize vector 
for(i in 1:nrow(dataFrame)){
  allRows[i] <- all(is.na(dataFrame[i,])) # store each output into allRows
}

blankRows <- which(allRows) # outputs indices of rows that contain "TRUE" (rows with all NA's) 
dataFrame <- dataFrame[-blankRows,] # remove those blankRows from dataFrame
```

Alternatively, you can use apply() to iterate through all rows. You can use this for a single data frame or a list of multiple data frames using a for loop. 
```{r eval=F}
# outputs indices of rows with all NA's
blankRows <- which(apply(dataFrame,1,function(x)all(is.na(x)))) 
```

##Columns
```{r eval=F}

i=0
allCols <- c()
for(i in 1:length(dataFrame)){ # length(dataFrame) gives us # of cols
  allRows[i] <- all(is.na(dataFrame[,i])) # notice that we switch where the i goes 
}

blankCols <- which(allCols)
dataFrame <- dataFrame[,-blankCols]
```

Alternatively:
```{r eval=F}
blankCols <- which(apply(dataFrame,2,function(x)all(is.na(x))))
```


#Searching Through Strings - Dates

Use the grepl() function to search for a particular string. Since we often have to reformat dates in our data sets, searching for particular dates or times could be useful. 

```{r}
# an example of common date/time scenarios
# this is usually a column within a data frame
dates <- c("3/4/2016", "3/4/16", "3-4-2016", "3-4-16","3-4-16 12:30",
           "3/4/2016", "3/4/16", "3-4-2016", "3-4-16","3-4-16 12:30",
           "3/4/2016", "3/4/16", "3-4-2016", "3-4-16","3-4-16 12:30")
```

Run unique() to see what kind of formats there are. 
```{r}
unique(dates)
```
The international standard format for dates and time are YYYY-MM-DD and hh:mm:ss respectively, while the combined date-time standard is YYYY-MM-DDThh:mm:ss. Often times, researchers' data contain dates and times in varying formats because it may have been inputted by different people. 

None of these are in the standard format, so we'll have to do some reformatting. 


The following code gives us the indices that contain "/2016".
```{r}
indDates <- which(grepl("/2016",dates))
indDates
```

Use as.POSIXct() to specify what our original date format is. 
Use format() to specify the format that we want.
Store values back into dates vector. 
```{r}
dates[indDates] <- format(as.POSIXct(dates[indDates], tz = "", format="%m/%d/%Y"), format = "%Y-%m-%d")
dates[indDates]
```
This same process works for all of the formats in our dates vector.
Note: 
"-16" is ambiguous because it could also refer to the day within an already standard-formatted date (e.g. 2018-05-16). Always check to make sure. 
We will reformat combined date/time items before the observations that don't contain times because they also contain "-16", which is ambiguous.

```{r}
indDates1 <- which(grepl("/16",dates))
dates[indDates1] <- format(as.POSIXct(dates[indDates1], tz = "", format="%m/%d/%y"), format = "%Y-%m-%d")

indDates2 <- which(grepl("-2016",dates))
dates[indDates2] <- format(as.POSIXct(dates[indDates2], tz = "", format="%m-%d-%Y"), format = "%Y-%m-%d")

indDates3 <- which(grepl("-16 ",dates))
dates[indDates3] <- format(as.POSIXct(dates[indDates3], tz = "", format="%m-%d-%y %H:%M"), format = "%Y-%m-%dT%H:%M:%S")

indDates4 <- which(grepl("-16",dates))
dates[indDates4] <- format(as.POSIXct(dates[indDates4], tz = "", format="%m-%d-%y"), format = "%Y-%m-%d")
```

Our final dates vector now looks like this: 
```{r echo=FALSE}
dates
```